---
title: "170111 Analysis of E Cell Antag Screen"
output: pdf_document
---

This is the third attempt to run a antagonist screen with the newly optimized E cells to make sure that H3A cells and E Cells show similar responses to odors. This run was different becuase a new aloquat of TMA was made and the antag odor block was diluted from the stock (instead of using an older dilution). In addition many extra TMA alone (or forskolin alone) wells were run throughout the plate (specifically in all of column 12 and all of row J). This will help us observe any plate effects that may be influencing these experiments. 

```{r, echo = FALSE, message=FALSE, include=FALSE}
#enter libraries
library(tidyverse)
library(drc)
library(platetools)
library(ggthemes)
library(viridis)

#import data
data <- read.csv("/Volumes/mainland/Projects/TAARs/E\ cell\ Optimization/AntagScreenRepeat_mini/170111_Results.csv", stringsAsFactors = FALSE)
#get rid of unnecessary columns
data <- subset(data, select = c("TransfectionDate", "PlateNum", "CloneKey", "Odor1K", "Odor2K","Odor1C", "Odor2C", "PlateLocation", "LucData", "RLData", "Notes","LabNoteBook", "TransfectionType", "CellType"))
#cuts down more columns to the useful ones for this analysis
data.sub <- subset(data, select = c("PlateNum", "CloneKey", "Odor1K", "Odor2K","Odor1C", "Odor2C", "PlateLocation", "LucData", "RLData", "CellType", "TransfectionType"))
data.sub$norm <- data.sub$LucData/data.sub$RLData #calculate normalized luc response
eCells <- subset(data.sub, CellType == "E Cells") #creates df with E cell data
H3A <- subset(data.sub, CellType == "H3A") #subsets df with H3A data

#####ANALYZE THE ANTAG RESPONSE: E Cells####
#E Cells
x = 817 #agonist #
y = "TMA" #agonist name
z = "hTAAR5" #receptor

Screen<-subset(eCells, Odor1K == x ) # get subset for the plate we are interested in
LowRL<-subset(Screen, Screen$RLData < 1000) #how many have low RL? (for H3A cells)
Screen<-subset(Screen, Screen$RLData > 1000) #subset out low RL values

#for E Cells, there are no RL values that low
#find standard deviation of the RL values for E Cells
meanRL <- mean(Screen$RLData)
sdRL <- sd(Screen$RLData)
#set min and max for more than 2 sd's outside of the mean RL value
minRL <- meanRL - 2*sdRL
maxRL <- meanRL + 2*sdRL
LowRL <- subset(Screen, Screen$RLData < minRL)
#no low RL for this set of E Cells
HighRL <- subset(Screen, Screen$RLData > maxRL)
#there are some values for highRL, but we are going to keep them for now. 

#Antag<-subset(Screen, Odor1InstockKey == x) ##subset fof agonist; we don't have TMA+TMA in this run
Agx2 <- subset(Screen, Odor2K == x) ##Agonist + Agonist in screen
Agx1 <- subset(Screen, Odor2K == 811) ##Agonist only
Screen$PerAg <- Screen$norm/(mean(Agx1$norm))*100 ##% of avg of Agonist only
Screen$PerInhib <- 100-Screen$PerAg ##%inhib
Antag <- Screen[order(Screen$PerAg) , ]
#add rankings to antag
Antag$rankInhib <- c(1:length(Antag$norm))
#pull out the "top hits" from old run
OldTopHits.E <- Antag[Antag$Odor2K %in% c(678, 486, 1221, 527, 821, 827, 817),]

```
Inhibition rankings of trans-2-nonen-1-ol (821) and Linalool (827, control odor).  
```{r, echo = FALSE}
OldTopHits.E #does this show them on the markdown file? 

#####Graph E Cell results#####
ggplot(data = Antag, aes(x = rankInhib, y = PerAg)) +
  geom_point() + 
  geom_point(data = subset(Antag, Odor2K == 811), aes(colour = "TMA Alone")) +
  ylab("% Response of mean 'agonist alone' response") +
  xlab("Rank of % inhibition") +
  ggtitle("E Cells with hTAAR5 Antagonist Response") +
  theme(legend.position = c(.1,.9),
        legend.title = element_blank(),
        panel.background = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = "black"))
```
  
The agonist alone control (TMA alone) is not at all consistent (pink dots). 
Let us look at the H3A plate with hTAAR5 and TMA to see if we observe a similar effect. 
```{r, echo=FALSE}
###Analysis for H3A Cells#####
x = 817
Screen<-subset(H3A, Odor1K == 817)
LowRL<-subset(Screen, Screen$RL < 1000) #how many have low RL?
#Screen<-subset(Screen, Screen$RL > 1000) #subset out low RL values

#Antag<-subset(Screen, Odor1InstockKey == x) ##subset fof agonist
Agx2 <- subset(Screen, Odor2K == x) ##Agonist + Agonist in screen
Agx1 <- subset(Screen, Odor2K == 811) ##Agonist only
Screen$PerAg <- Screen$norm/(mean(Agx1$norm))*100 ##% of avg of Agonist only
Screen$PerInhib <- 100-Screen$PerAg ##%inhib
Antag <- Screen[order(Screen$PerAg) , ]
#add rankings to antag
Antag$rankInhib <- c(1:length(Antag$norm))
#pull out the "top hits"
OldTopHits.H3A <- Antag[Antag$Odor2K %in% c(678, 486, 1221, 527, 821, 827, 817),]

```
%Inhibition rankings of trans-2-nonen-1-ol (821) and the control Linalool(827)  
```{r, echo = FALSE}
OldTopHits.H3A
#####Graph E Cell results#####
ggplot(data = Antag, aes(x = rankInhib, y = PerAg)) +
  geom_point() + 
  geom_point(data = subset(Antag, Odor2K == 811), aes(colour = "TMA Alone")) +
  ylab("% Response of mean 'agonist alone' response") +
  xlab("Rank of % inhibition") +
  ggtitle("H3A cells with hTAAR5 Antagonist Response") +
  theme(legend.position = c(.1,.9),
        legend.title = element_blank(),
        panel.background = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = "black"))

```
  
The H3A TMA responses are also very inconsistent. It is likely that we are seeing some sort of plate effect. 

#Examining Plate Effects
```{r, echo = FALSE}
##heat maps of the  plate effects
raw_map(data = eCells[which(eCells$Odor1K == 817),]$norm,
      well =eCells[which(eCells$Odor1K == 817),]$PlateLocation, 
      plate =384) +
  theme_dark() +
  scale_fill_viridis() +
  ggtitle("Normalized luciferase values of E-Cells with hTAAR5")

raw_map(data = eCells[which(eCells$Odor1K == 77),]$norm,
      well =eCells[which(eCells$Odor1K == 77),]$PlateLocation, 
      plate =384) +
  theme_dark() +
  scale_fill_viridis() +
  ggtitle("Normalized luciferase values of E-Cells with Rho")

#H3A
raw_map(data = H3A[which(H3A$Odor1K == 817),]$norm,
      well =H3A[which(H3A$Odor1K == 817),]$PlateLocation, 
      plate =384) +
  theme_dark() +
  scale_fill_viridis() +
  ggtitle("Normalized luciferase values of H3A with hTAAR5")

raw_map(data = H3A[which(H3A$Odor1K == 77),]$norm,
      well =H3A[which(H3A$Odor1K == 77),]$PlateLocation, 
      plate =384) +
  theme_dark() +
  scale_fill_viridis() +
  ggtitle("Normalized luciferase values of H3A with Rho")
```
  
The heat map of E cells with hTAAR5 (top) shows a very clear gradient with increasing activation down the plate from A to P. This is especially evident when you focus on the TMA alone in column 12. 
The other plates do not show as obvious a gradient.
Since this is looking at normalized values, I will next examine whether the luciferase or RL values are the cause of this gradient.  

```{r, echo = FALSE}
raw_map(data = eCells[which(eCells$Odor1K == 817),]$LucData,
      well =eCells[which(eCells$Odor1K == 817),]$PlateLocation, 
      plate =384) +
  theme_dark() +
  scale_fill_viridis() +
  ggtitle("Luciferase values of E-Cells with hTAAR5 (not normalized!)")
```  
  
The luciferase does not show any obvious pattern or gradient.  
```{r, echo = FALSE}
raw_map(data = eCells[which(eCells$Odor1K == 817),]$RLData,
      well =eCells[which(eCells$Odor1K == 817),]$PlateLocation, 
      plate =384) +
  theme_dark() +
  scale_fill_viridis() +
  ggtitle("RL values of E-Cells with hTAAR5 (not normalized!)")
``` 
  
RL values show a clear gradient with high reads on the top of the plate to lower reads on the bottom (because normalized data divides by RL, this is why we see the opposite gradient on the normalized plate).  
  
Interestingly, and frightenly, we see the same RL gradient on each of the other plates of this run.  
```{r, echo = FALSE}
raw_map(data = eCells[which(eCells$Odor1K == 77),]$RLData,
      well =eCells[which(eCells$Odor1K == 77),]$PlateLocation, 
      plate =384) +
  theme_dark() +
  scale_fill_viridis() +
  ggtitle("RL values of E-Cells with Rho")

raw_map(data = H3A[which(H3A$Odor1K == 817),]$RLData,
      well =H3A[which(H3A$Odor1K == 817),]$PlateLocation, 
      plate =384) +
  theme_dark() +
  scale_fill_viridis() +
  ggtitle("RL values of H3A with hTAAR5")

raw_map(data = H3A[which(H3A$Odor1K == 817),]$RLData,
      well =H3A[which(H3A$Odor1K == 817),]$PlateLocation, 
      plate =384) +
  theme_dark() +
  scale_fill_viridis() +
  ggtitle("RL values of H3A with Rho")
``` 
  
Why are we seeing this gradient of RL, and is it on other run of 384 plates?  
  
To examine this further, I will look at the last two antag screens and see if the RL values show the same gradient. 

#Examine previous Antag screens for RL gradient  
  
For the screen on 11-4-16  
```{r, echo = FALSE}
#analyze the data from 9-30-16
#this was an antagonist screen comparing results between H3A and E cells, and to make sure that the same top hits came up as the last time we ran the screen. 
#import data
data <- read.csv("/Volumes/mainland/Projects/TAARs/E\ cell\ Optimization/AntagScreenRepeat_mini/161104_Results.csv", stringsAsFactors = FALSE)
#get rid of unnecessary columns
#gives all data
data <- subset(data, select = c("TransfectionDate", "PlateNum", "CloneKey", "Odor1K", "Odor2K","Odor1C", "Odor2C", "PlateLocation", "LucData", "RLData", "Notes","LabNoteBook", "TransfectionType", "CellType"))
#cuts down more columns to the useful ones for this analysis
data.sub <- subset(data, select = c("PlateNum", "CloneKey", "Odor1K", "Odor2K","Odor1C", "Odor2C", "PlateLocation", "LucData", "RLData", "CellType"))

#calculate normalized luc response
data.sub$norm <- data.sub$LucData/data.sub$RLData
#combine per cell type
eCells <- subset(data.sub, PlateNum == 1 | PlateNum == 2)
H3A <- subset(data.sub, PlateNum == 3 | PlateNum == 4)

raw_map(data = eCells[which(eCells$Odor1K == 817),]$RLData,
      well =eCells[which(eCells$Odor1K == 817),]$PlateLocation, 
      plate =384) +
  theme_dark() +
  scale_fill_viridis() +
  ggtitle("RL values of eCells with hTAAR5")

raw_map(data = eCells[which(eCells$Odor1K == 77),]$RLData,
      well =eCells[which(eCells$Odor1K == 77),]$PlateLocation, 
      plate =384) +
  theme_dark() +
  scale_fill_viridis() +
  ggtitle("RL values of eCells with Rho")

raw_map(data = H3A[which(H3A$Odor1K == 817),]$RLData,
      well =H3A[which(H3A$Odor1K == 817),]$PlateLocation, 
      plate =384) +
  theme_dark() +
  scale_fill_viridis() +
  ggtitle("RL values of H3A with hTAAR5")

raw_map(data = H3A[which(H3A$Odor1K == 77),]$RLData,
      well =H3A[which(H3A$Odor1K == 77),]$PlateLocation, 
      plate =384) +
  theme_dark() +
  scale_fill_viridis() +
  ggtitle("RL values of H3A with Rho")

```
  
All of the plates from the run on 11-4-16 also show this RL pattern across the plate. What could be causing this?  
Additionally, on the Ecell hTAAR5 plate, there is a clear pipetting effect (every other row, which is the way cells are plated and transfected), but this is the only place where this has turned up.  
  
Examine Run from 10/1/16 for RL pattern  
```{r, echo = FALSE}
plate1 <- read.csv("/Volumes/mainland/Projects/TAARs/E\ cell\ Optimization/AntagScreenRepeat_mini/160930_plate1.csv", stringsAsFactors = FALSE)[,1:11]
plate2 <- read.csv("/Volumes/mainland/Projects/TAARs/E\ cell\ Optimization/AntagScreenRepeat_mini/160930_plate2.csv", stringsAsFactors = FALSE)[,1:11]
plate3 <- read.csv("/Volumes/mainland/Projects/TAARs/E\ cell\ Optimization/AntagScreenRepeat_mini/160930_plate3.csv", stringsAsFactors = FALSE)[,1:11]
plate4 <- read.csv("/Volumes/mainland/Projects/TAARs/E\ cell\ Optimization/AntagScreenRepeat_mini/160930_plate4.csv", stringsAsFactors = FALSE)[,1:11]
#combine per cell type
eCells <- rbind(plate1, plate2)
H3A <- rbind(plate3, plate4)
#calculate the normalized response
eCells$norm <- eCells$Luc/eCells$RL
H3A$norm <- H3A$Luc/H3A$RL

raw_map(data = eCells[which(eCells$agonistInStockKey == 817),]$RL,
      well =eCells[which(eCells$agonistInStockKey == 817),]$plateLocation, 
      plate =384) +
  theme_dark() +
  scale_fill_viridis() +
  ggtitle("RL values of eCells with hTAAR5")

raw_map(data = eCells[which(eCells$agonistInStockKey == 77),]$RL,
      well =eCells[which(eCells$agonistInStockKey == 77),]$plateLocation, 
      plate =384) +
  theme_dark() +
  scale_fill_viridis() +
  ggtitle("RL values of eCells with Rho")

raw_map(data = H3A[which(H3A$agonistInStockKey == 817),]$RL,
      well =H3A[which(H3A$agonistInStockKey == 817),]$plateLocation, 
      plate =384) +
  theme_dark() +
  scale_fill_viridis() +
  ggtitle("RL values of H3A with hTAAR5")

raw_map(data = H3A[which(H3A$agonistInStockKey == 77),]$RL,
      well =H3A[which(H3A$agonistInStockKey == 77),]$plateLocation, 
      plate =384) +
  theme_dark() +
  scale_fill_viridis() +
  ggtitle("RL values of H3A with Rho")
```
  
All of the plates from the experiment on 10-4-16 show the same RL patterns as well.  
  
This explains the unusual result I have been seeing in these runs. In all previous experiments, the TMA alone has had the lowest normalized luciferase value, even compared to potential antagonists. The TMA alone wells have always been at the top of the plate, meaning that becuase RL is highest there, the normalized value is lowest.  
  
Why does the RL show this gradient? should I analyze these plates based on luciferase alone? Should I find a way to normalize the plates that takes the RL gradient into account? 
