---
title: "170111 Analysis of E Cell Antag Screen"
output: html_document
---

This is the third attempt to run a antagonist screen with the newly optimized E cells to make sure that H3A cells and E Cells show similar responses to odors. This run was different becuase a new aloquat of TMA was made and the antag odor block was diluted from the stock (instead of using an older dilution). In addition many extra TMA alone (or forskolin alone) wells were run throughout the plate (specifically in all of column 12 and all of row J). This will help us observe any plate effects that may be influencing these experiments. 

```{r, echo = FALSE, message=FALSE, include=FALSE}
#enter libraries
library(tidyverse)
library(drc)
library(platetools)
library(ggthemes)
library(viridis)

#import data
data <- read.csv("/Volumes/mainland/Projects/TAARs/E\ cell\ Optimization/AntagScreenRepeat_mini/170111_Results.csv", stringsAsFactors = FALSE)
#get rid of unnecessary columns
data <- subset(data, select = c("TransfectionDate", "PlateNum", "CloneKey", "Odor1K", "Odor2K","Odor1C", "Odor2C", "PlateLocation", "LucData", "RLData", "Notes","LabNoteBook", "TransfectionType", "CellType"))
#cuts down more columns to the useful ones for this analysis
data.sub <- subset(data, select = c("PlateNum", "CloneKey", "Odor1K", "Odor2K","Odor1C", "Odor2C", "PlateLocation", "LucData", "RLData", "CellType", "TransfectionType"))
data.sub$norm <- data.sub$LucData/data.sub$RLData #calculate normalized luc response
eCells <- subset(data.sub, CellType == "E Cells") #creates df with E cell data
H3A <- subset(data.sub, CellType == "H3A") #subsets df with H3A data

#####ANALYZE THE ANTAG RESPONSE: E Cells####
#E Cells
x = 817 #agonist #
y = "TMA" #agonist name
z = "hTAAR5" #receptor

Screen<-subset(eCells, Odor1K == x ) # get subset for the plate we are interested in
LowRL<-subset(Screen, Screen$RLData < 1000) #how many have low RL? (for H3A cells)
Screen<-subset(Screen, Screen$RLData > 1000) #subset out low RL values

#for E Cells, there are no RL values that low
#find standard deviation of the RL values for E Cells
meanRL <- mean(Screen$RLData)
sdRL <- sd(Screen$RLData)
#set min and max for more than 2 sd's outside of the mean RL value
minRL <- meanRL - 2*sdRL
maxRL <- meanRL + 2*sdRL
LowRL <- subset(Screen, Screen$RLData < minRL)
#no low RL for this set of E Cells
HighRL <- subset(Screen, Screen$RLData > maxRL)
#there are some values for highRL, but we are going to keep them for now. 

#Antag<-subset(Screen, Odor1InstockKey == x) ##subset fof agonist; we don't have TMA+TMA in this run
Agx2 <- subset(Screen, Odor2K == x) ##Agonist + Agonist in screen
Agx1 <- subset(Screen, Odor2K == 811) ##Agonist only
Screen$PerAg <- Screen$norm/(mean(Agx1$norm))*100 ##% of avg of Agonist only
Screen$PerInhib <- 100-Screen$PerAg ##%inhib
Antag <- Screen[order(Screen$PerAg) , ]
#add rankings to antag
Antag$rankInhib <- c(1:length(Antag$norm))
#pull out the "top hits" from old run
OldTopHits.E <- Antag[Antag$Odor2K %in% c(678, 486, 1221, 527, 821, 827, 817),]

```
Inhibition rankings of trans-2-nonen-1-ol (821) and Linalool (827, control odor).  
```{r, echo = FALSE}
OldTopHits.E #does this show them on the markdown file? 

#####Graph E Cell results#####
ggplot(data = Antag, aes(x = rankInhib, y = PerAg)) +
  geom_point() + 
  geom_point(data = subset(Antag, Odor2K == 811), aes(colour = "TMA Alone")) +
  ylab("% Response of mean 'agonist alone' response") +
  xlab("Rank of % inhibition") +
  ggtitle("E Cells with hTAAR5 Antagonist Response") +
  theme(legend.position = c(.1,.9),
        legend.title = element_blank(),
        panel.background = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = "black"))
```
  
The agonist alone control (TMA alone) is not at all consistent (pink dots). 
Let us look at the H3A plate with hTAAR5 and TMA to see if we observe a similar effect. 
```{r, echo=FALSE}
###Analysis for H3A Cells#####
x = 817
Screen<-subset(H3A, Odor1K == 817)
LowRL<-subset(Screen, Screen$RL < 1000) #how many have low RL?
#Screen<-subset(Screen, Screen$RL > 1000) #subset out low RL values

#Antag<-subset(Screen, Odor1InstockKey == x) ##subset fof agonist
Agx2 <- subset(Screen, Odor2K == x) ##Agonist + Agonist in screen
Agx1 <- subset(Screen, Odor2K == 811) ##Agonist only
Screen$PerAg <- Screen$norm/(mean(Agx1$norm))*100 ##% of avg of Agonist only
Screen$PerInhib <- 100-Screen$PerAg ##%inhib
Antag <- Screen[order(Screen$PerAg) , ]
#add rankings to antag
Antag$rankInhib <- c(1:length(Antag$norm))
#pull out the "top hits"
OldTopHits.H3A <- Antag[Antag$Odor2K %in% c(678, 486, 1221, 527, 821, 827, 817),]
```
%Inhibition rankings of trans-2-nonen-1-ol (821) and the control Linalool(827)  
```{r, echo = FALSE}
OldTopHits.H3A
#####Graph E Cell results#####
ggplot(data = Antag, aes(x = rankInhib, y = PerAg)) +
  geom_point() + 
  geom_point(data = subset(Antag, Odor2K == 811), aes(colour = "TMA Alone")) +
  ylab("% Response of mean 'agonist alone' response") +
  xlab("Rank of % inhibition") +
  ggtitle("H3A cells with hTAAR5 Antagonist Response") +
  theme(legend.position = c(.1,.9),
        legend.title = element_blank(),
        panel.background = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = "black"))

```
  
The H3A TMA responses are also very inconsistent. It is likely that we are seeing some sort of plate effect. 

#Examining Plate Effects
```{r, echo = FALSE}
##heat maps of the 
raw_map(data = Screen$LucData,
      well = Screen$PlateLocation, 
      plate =384) +
  theme_dark() +
  scale_fill_viridis() +
  ggtitle("Luc values of E-Cells and hTAAR5")

raw_map(data = eCells$norm,
        well = eCells$PlateLocation, 
        plate =384, 
        plate_id = eCells$PlateNum) +
  theme_dark() +
  scale_fill_viridis() +
  ggtitle("normalized luciferase values of E Cells")
```
